---
title: "Applied Bayesian Analysis : NCSU ST 540"
subtitle: "Midterm2"
author: "Bruce Campbell"
fontsize: 11pt
output: pdf_document
bibliography: BruceCampbell_ST540_HW_1.bib
---

---
```{r setup, include=FALSE,echo=FALSE}
rm(list = ls())
setwd("d:/brucebcampbell-git/bayesian-learning-with-R")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=5)
knitr::opts_chunk$set(fig.width=6)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_knit$set(root.dir = ".")
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=38),tidy=TRUE)
library(latex2exp)   
library(pander)
library(ggplot2)
library(GGally)
```



# VAR(1) in JAGS

This section is a test section where we generate and fit a vector autoregressive  model - $VAR(1) \in \mathbf{R}^6$ given by 

$$y_{t} = \nu + \rho * y_{t-1} + \epsilon$$ 

$$\epsilon \sim N(0,\Sigma)$$

We use the $y1$ data to calculate a NaN firendly sample covariance and then we find the nearest positive semidefinite matrix to use to generate data for the model. 

Notes
- imputation is not working for this model
- this section seeks to find the parameters that best explain the data

```{r}
library(rjags)
library(coda)
library(modeest)
library(MASS)
load("E2.RData")

DEBUG <- TRUE
if(DEBUG)
{
nSamples <- 5000
n.chains <- 1
} else
{
nSamples <- 40000
n.chains <- 8
}
cor.y1 <- cor(Y1, use = "pairwise.complete.obs")
cov.y1  <- cov(Y1, use = "pairwise.complete.obs")

N <- nrow(Y1)
p = 6
Y1.scaled <- scale(Y1)
# Try to simulate using the corr
#install.packages("Matrix")
library("Matrix")

sig <- nearPD(cov.y1)
N = 365
Sigma = sig$mat 
rho = .6
nu = matrix(rep(.1,6), p, 1)
y = matrix(NA, N,p)
y[1,] = nu
for(t in 2:N) 
{
  y[t,] = mvrnorm(1, nu + rho * y[t-1,], Sigma)
}
plot(y[,1])
# Jags code to fit the model to the simulated data
model_code = '
model
{
  # Likelihood
  for (t in 2:N) 
  {
    y[t, ] ~ dmnorm(mu[t, ], precisionAR)
    mu[t, 1:p] <- nu + rho * y[t-1,]
  }
  precisionAR ~ dwish(I, p+1)
  Sigma <- inverse(precisionAR)  
  
  # Priors
  rho ~ dunif(-1, 1)
  for(i in 1:p) 
  {
    nu[i] ~ dnorm(0, 0.01)
  }

  # Missing data model for y - not working in JAGS
  #for(i in 1:N)
  #{
    #y[i,1:p]~dmnorm(x_mn[],x_prec[,])
  #}
  
  # Priors for missing-data model parameters
  for(j in 1:p)
  {
    x_mn[j]~dnorm(0,0.01)
  }
  x_prec[1:p,1:p]~dwish(R[,],k)
  x_cov[1:p,1:p]<-inverse(x_prec[,])
  k <- p+0.1
  for(j1 in 1:p)
  {
    for(j2 in 1:p)
    {
      R[j1,j2] <- 0.1*equals(j1,j2)
    }     
  }
}
'
# Set up the data
model_data = list(N = N, p = p, y = y, I = diag(p))
# Choose the parameters to watch
model_parameters =  c("nu", "rho", "Sigma","y")

model <- jags.model(textConnection(model_code),data = model_data,n.chains = n.chains)#Compile Model Graph

update(model, nSamples, progress.bar="none"); # Burnin
out.coda  <- coda.samples(model, variable.names=model_parameters,n.iter=2*nSamples) 

#plot(out.coda )

if(n.chains > 1)
{
  gelman.srf <-gelman.diag(out.coda)
  plot(gelman.srf$psrf,main = "Gelman Diagnostic")
}

chains.ess <- lapply(out.coda,effectiveSize)
first.chain.ess <- chains.ess[1]
plot(unlist(first.chain.ess), main="Effective Sample Size")

chain <- out.coda[[1]]
posterior.means <- list()
posterior.modes <- list()
for( i in 1:length(colnames(chain)) )
{  
  colname <- colnames(chain)[i]
  samples <- chain[,i]
  posterior.means[colname] <-mean(samples)
  posterior.modes[colname] <-mlv(samples)$M
}

# plot(posterior.means,posterior.modes) # Nice and unimodal

Sigma.map <- unlist(posterior.means)[1:36]
Sigma.model<- unlist(sig$mat)
plot(Sigma.map,Sigma.model)
rho.map <- unlist(posterior.means)[43]

nu.map <-  unlist(posterior.means)[37:42]
 
# y.map <-  matrix(unlist(posterior.means)[44:2233],ncol=6, byrow=FALSE)
# plot(y[,1]-y.map[,1]) 
```

# Missing data imputation in Openbugs

Here we try to implement missing data imputation in OpenBugs. We use the data set y1

```{r, results='hide', message=FALSE, warning=FALSE}
rm(list = ls())
setwd("d:/brucebcampbell-git/bayesian-learning-with-R")
load("E2.RData")
library(R2OpenBUGS)
library(rjags)
library(coda)
library(modeest)
N <- nrow(Y1)
p = 6
x <- scale(Y1)

mlr_model2 <- function(){
  for(i in 1:N)
  {
    theta[i,1:p] ~ dmnorm(x[i,1:p] ,precision2[,])
  }
  
  # Prior for likelihood parameters: mu2, precision2, rho
  rho  ~  dunif(-1,1)
  
  for(j in 1:p)
  {
    mu2[j]  ~  dnorm(0,0.01)
  }
  
  precision2[1:p,1:p] ~ dwish(R[,],k)
  
  # Missing data model for x
  for(i in 1:N){
    x[i,1:p]~dmnorm(x_mn[],x_prec[,])
  }
  
  # Priors for missing-data model parameters
  for(j in 1:p){
    x_mn[j]~dnorm(0,0.01)
  }
  x_prec[1:p,1:p]~dwish(R[,],k)
  x_cov[1:p,1:p]<-inverse(x_prec[,])
  
  k <- p+0.1
  for(j1 in 1:p)
  {
    for(j2 in 1:p)
    {
      R[j1,j2] <- 0.1*equals(j1,j2)
    }
  }
}

n.chains = 1
nSamples=10000
stacks_dat <- list(x=x,p = 6,   N = 365)
mlr_inits <- function() {   list( rho = 0.00) }

samps <- bugs(data = stacks_dat, 
              inits = mlr_inits, 
              parameters.to.save = c("theta"), 
              model.file = mlr_model2, 
              codaPkg = TRUE,
              n.chains = n.chains, n.burnin=2000, n.iter = nSamples, n.thin=10, DIC=F)

out.coda <- read.bugs(samps) 

if(n.chains > 1)
{
  gelman.srf <-gelman.diag(out.coda)
  count.coeff.gt <- sum(gelman.srf$psrf>1.1)
  count.coeff.gt
}

chains.ess <- lapply(out.coda,effectiveSize)

first.chain.ess <- chains.ess[1]
plot(unlist(first.chain.ess), main="Effective Sample Size")

chain <- out.coda[[1]]
posterior.means <- list()
posterior.modes <- list()

for( i in 1:(365*6) )
{  
  colname <- colnames(chain)[i]
  
  samples <- chain[,i]
  
  posterior.means[i] <-mean(samples)
  
  posterior.modes[i] <-mlv(samples)$M
}

plot(posterior.means, posterior.modes)

theta.map <-  matrix(unlist(posterior.means)[1:2190],ncol=6, byrow=FALSE)

unscaled.theta.map <- ( theta.map +colMeans(Y1,na.rm = TRUE)) *  apply(Y1, 2,sd,na.rm = TRUE)

write.csv(unscaled.theta.map,file = "unscaled-theta-map.csv")
hist(unscaled.theta.map[,1])

hist(Y1[,1])


```