---
title: "Applied Bayesian Analysis : NCSU ST 540"
subtitle: "Midterm2"
author: "Bruce Campbell"
fontsize: 11pt
output: pdf_document
bibliography: BruceCampbell_ST540_HW_1.bib
---

---
```{r setup, include=FALSE,echo=FALSE}
rm(list = ls())
setwd("c:/e/brucebcampbell-git/bayesian-learning-with-R")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=5)
knitr::opts_chunk$set(fig.width=6)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_knit$set(root.dir = ".")
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=38),tidy=TRUE)
library(latex2exp)   
library(pander)
library(ggplot2)
library(GGally)
```

This section is a test section where we generate and fit a vector autoregressive  model - $VAR(1) \in \mathbf{R}^6$ given by 

$$y_{t} = \nu + \rho * y_{t-1} + \epsilon$$ 

$$\epsilon \sim N(0,\Sigma)$$

We use the $y1$ data to calculate a NaN firendly sample covariance and then we find the nearest positive semidefinite matrix to use to generate data for the model. 

Notes
- imputation is not working for this model
- this section seeks to find the parameters that best explain the data

```{r}
library(MASS)
library("Matrix")
library(rjags)
library(coda)
library(modeest)
library(MASS)
load("E2.RData")
cov.y1  <- cov(Y1, use = "pairwise.complete.obs")
heatmap(cov.y1,main = "cov of Y1")
sig <- nearPD(cov.y1)
heatmap(as.matrix(sig$mat),main = "cov of y - our test data ")
Sigma = sig$mat
cov.y1  <- cov(Y1, use = "pairwise.complete.obs")
ggplot(data.frame(Y3=Y3) ,aes(x=1:365, y=Y3)) + geom_point(alpha=0.25) + geom_smooth( method="loess", span=0.22) +ggtitle("Y3")
N <- nrow(Y1)
p = 6

# Jags code to fit the model to the simulated data
model_code = '
model
{
  Y1pred[1,1:p] ~ dmnorm( Y1[1,1:p] ,precision[,])
  
  for(i in 2:N) {
    Y1pred[i,1:p] ~ dmnorm( theta[i,1:p] ,precision[,])
    
    #Y2[i] ~ dnorm(Y1[i-1,1],sigmaY2)
    
    theta[i,1]<-  mu2[1] + rho * Y1[i-1,1]
    theta[i,2]<-  mu2[2] + rho * Y1[i-1,2]
    theta[i,3]<-  mu2[3] + rho * Y1[i-1,3]
    theta[i,4]<-  mu2[4] + rho * Y1[i-1,4]
    theta[i,5]<-  mu2[5] + rho * Y1[i-1,5]
    theta[i,6]<-  mu2[6] + rho * Y1[i-1,6]

    #Y3[i] ~ dnorm(thetaBar[i],sigmaY3)
    #thetaBar[i] <- 1/6 *(theta[i,2] +theta[i,2]+ theta[i,3]+ theta[i,4] +theta[i,5] +theta[i,6])
    }
  
  # Priors
  rho  ~  dunif(0,1)
  sigmaY3 ~ dnorm(0, 0.01)
  sigmaY2  ~ dnorm(0, 0.01)
    
  for(j in 1:p)
  {
    mu2[j]  ~  dnorm(0,0.01)
  }
  
  precision[1:p,1:p] ~ dwish(R[,],k)
  
  SigmaVAR[1:p,1:p] <- inverse(precision[1:p,1:p])
  
  # Missing data model for x
  for(i in 1:N){
    Y1[i,1:p]~dmnorm(Y1_mn[],Y1_prec[,])
  }
  
  # Priors for missing-data model parameters
  for(j in 1:p){
    Y1_mn[j]~dnorm(0,0.01)
  }
  Y1_prec[1:p,1:p]~dwish(R[,],k)
  
  k <- p+0.1
  for(j1 in 1:p)
  {
    for(j2 in 1:p)
    {
      R[j1,j2] <- 0.1*equals(j1,j2)
    }
  }
}
'
cm <-colMeans(Y1.aug,na.rm = TRUE)

Y1_1.mod <- Y1[,1]
#Y1_1.mod <- (Y1[,1] + Y2) /2
naidx <- is.na(Y1_1.mod)
Y1_1.mod[naidx] <- Y2[naidx]
Y1.aug[,1] <- Y1_1.mod

for(i in 1:N)
{
  for(j in 1:p)
  {
    is.na(Y1.aug[i,j])
    if(is.na(Y1.aug[i,j]))
    {
      if(is.na(Y3[i]))
      {
        Y1.aug[i,j] <- cm[j]   
      }else
      {
        Y1.aug[i,j] = Y3[i]
      }
    }
  }
}
# Set up the data
model_data = list(Y1=scale(Y1.aug),Y2=as.vector(scale(Y2)),Y3=as.vector(scale(Y3)),N = N, p = p)
# Choose the parameters to watch
model_parameters =  c("theta","SigmaVAR" )

n.chains <- 2
model <- jags.model(textConnection(model_code),data = model_data,n.chains = n.chains)#Compile Model Graph

update(model, 5000, progress.bar="none"); # Burnin

out.coda  <- coda.samples(model, variable.names=model_parameters,n.iter=1000) 
save(out.coda,file = "out.coda_JAGS_VAR1.RData")


if(n.chains > 1)
{
  g <- matrix(NA, nrow=nvar(out.coda), ncol=2)
  for (v in 1:nvar(out.coda)) {
   g[v,] <- gelman.diag(out.coda[,v])$psrf
  }
  count.coeff.gt <- sum(g[,1]>1.1)
  count.coeff.gt
  plot(g[,1],main="Gelman-Rubin ")
}

chains.ess <- lapply(out.coda,effectiveSize)

first.chain.ess <- chains.ess[1]
plot(unlist(first.chain.ess), main="Effective Sample Size")

chain <- out.coda[[1]]

Y1.posterior.modes <- list()

for( i in 1:(365*6) )
{  
  colname <- colnames(chain)[i]
  if(grepl("Sigma",colname)  )
  {
    print(paste("skipping ",colname,sep=""))
    next
  }
  # samples <- chain[,i]
  # samplesId <- samples<4
  # red <- samples[samplesId]
  Y1.posterior.modes[colname] <-mlv(samples)$M
}

theta.map <-  matrix(unlist(Y1.posterior.modes),ncol=6, byrow=FALSE)

unscaled.theta.map <- ( theta.map +colMeans(Y1,na.rm = TRUE)) *  apply(Y1, 2,sd,na.rm = TRUE)

image(Y1, main="Y1")

image(unscaled.theta.map, main="Imputed")

SigmaVAR.posterior.modes <- list()

for( i in 1:(365*6+36) )
{  
  colname <- colnames(chain)[i]
  if(grepl("Sigma",colname) ==FALSE )
  {
    next
  }
  samples <- chain[,i]
  SigmaVAR.posterior.modes[colname] <-mlv(samples)$M
}

Sigma.map <-  matrix(unlist(SigmaVAR.posterior.modes),ncol=6, byrow=FALSE)
image(Sigma.map, main = "MAP Sigma")
image(cov.y1,main = "cov of Y1")

b <- unscaled.theta.map[!is.na(Y1)]
c <- Y1[!is.na(Y1)]

plot(b,c)
write.csv(unscaled.theta.map,file = "generativemodel - unscaled-theta-map.csv")

```







