---
title: "Applied Bayesian Analysis : NCSU ST 540"
subtitle: "Midterm2"
author: "Bruce Campbell"
fontsize: 11pt
output: pdf_document
bibliography: BruceCampbell_ST540_HW_1.bib
---

---
```{r setup, include=FALSE,echo=FALSE}
rm(list = ls())
setwd("c:/e/brucebcampbell-git/bayesian-learning-with-R")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(prompt=FALSE)
knitr::opts_chunk$set(fig.height=5)
knitr::opts_chunk$set(fig.width=6)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_knit$set(root.dir = ".")
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=38),tidy=TRUE)
library(latex2exp)   
library(pander)
library(ggplot2)
library(GGally)
```



# Test section - VAR(1) in JAGS

This section is a test section where we generate and fit a vector autoregressive  model - $VAR(1) \in \mathbf{R}^6$ given by 

$$y_{t} = \nu + \rho * y_{t-1} + \epsilon$$ 

$$\epsilon \sim N(0,\Sigma)$$

We use the $y1$ data to calculate a NaN firendly sample covariance and then we find the nearest positive semidefinite matrix to use to generate data for the model. 

Notes
- imputation is not working for this model
- this section seeks to find the parameters that best explain the data

```{r,echo=FALSE}
library(rjags)
library(coda)
library(modeest)
library(MASS)
load("E2.RData")
cor.y1 <- cor(Y1, use = "pairwise.complete.obs")
cov.y1  <- cov(Y1, use = "pairwise.complete.obs")
heatmap(cor.y1,main = "corr of Y1")
ggplot(data.frame(Y3=Y3) ,aes(x=1:365, y=Y3)) + geom_point(alpha=0.25) + geom_smooth( method="loess", span=0.22) +ggtitle("Y3")
ggplot(data.frame(mean.Y1=rowMeans(Y1,na.rm = TRUE)) ,aes(x=1:365, y=mean.Y1)) + geom_point(alpha=0.25) + geom_smooth( method="loess", span=0.22) +ggtitle("Row Means Y1")
N <- nrow(Y1)
p = 6
Y1.scaled <- scale(Y1)
# Try to simulate using the corr
#install.packages("Matrix")
library("Matrix")

sig <- nearPD(cov.y1)
N = 365
Sigma = sig$mat 
rho = .6
nu = matrix(rep(.1,6), p, 1)
y = matrix(NA, N,p)
y[1,] = nu
for(t in 2:N) 
{
  y[t,] = mvrnorm(1, nu + rho * y[t-1,], Sigma)
}

#Now Make Test data for validation of the Openbugs 
# y is our ground truth we add noise, calculate Y2 and Y3, and finally do the data redaction
y.test = matrix(NA, N,p)
v0 <- rep(0,6)
y.test[1,] = y[1,] +mvrnorm(1,v0, Sigma)
for(t in 2:N) 
{
  y.test[t,] = y[t,]+ mvrnorm(1, v0, Sigma*0.1)
}

Y1.test <-y
Y2.test <- y.test[,1]
Y3.test <- rowMeans(y.test)

#redaction pproportions
Y1.redaction.prop <- .8
Y2.redaction.prop <- .1
Y3.redaction.prop <- .1

Y1.test.redacted <- y.test
Y2.test.redacted <- Y2.test
Y3.test.redacted <- Y3.test

Y1.test.redacted <- unlist(Y1.test.redacted)
Y1.NA.index <- sample(length(Y1.test.redacted), floor(length(Y1.test.redacted)* Y1.redaction.prop))
Y1.test.redacted[Y1.NA.index] <- NA
Y1.test.redacted<-matrix(Y1.test.redacted,ncol=6, byrow=FALSE)

Y2.NA.index <- sample(length(Y2.test.redacted), floor(length(Y2.test.redacted)* Y2.redaction.prop))
Y2.test.redacted[Y2.NA.index] <- NA

Y3.NA.index <- sample(length(Y3.test.redacted), floor(length(Y3.test.redacted)* Y3.redaction.prop))
Y3.test.redacted[Y3.NA.index] <- NA
save(Y1.test,Y2.test,Y3.test,Y1.test.redacted,Y2.test.redacted,Y3.test.redacted,file="TestData.RData")


plot(y[,1])
# Jags code to fit the model to the simulated data
model_code = '
model
{
  # Likelihood
  for (t in 2:N) 
  {
    y[t, ] ~ dmnorm(mu[t, ], precisionAR)
    mu[t, 1:p] <- nu + rho * y[t-1,]
  }
  precisionAR ~ dwish(I, p+1)
  Sigma <- inverse(precisionAR)  
  
  # Priors
  rho ~ dunif(-1, 1)
  for(i in 1:p) 
  {
    nu[i] ~ dnorm(0, 0.01)
  }

  # Missing data model for y - not working in JAGS
  #for(i in 1:N)
  #{
    #y[i,1:p]~dmnorm(x_mn[],x_prec[,])
  #}
  
  # Priors for missing-data model parameters
  for(j in 1:p)
  {
    x_mn[j]~dnorm(0,0.01)
  }
  x_prec[1:p,1:p]~dwish(R[,],k)
  x_cov[1:p,1:p]<-inverse(x_prec[,])
  k <- p+0.1
  for(j1 in 1:p)
  {
    for(j2 in 1:p)
    {
      R[j1,j2] <- 0.1*equals(j1,j2)
    }     
  }
}
'
# Set up the data
model_data = list(N = N, p = p, y = y, I = diag(p))
# Choose the parameters to watch
model_parameters =  c("nu", "rho", "Sigma","y")

model <- jags.model(textConnection(model_code),data = model_data,n.chains = n.chains)#Compile Model Graph
update(model, nSamples, progress.bar="none"); # Burnin
if(FALSE)
{ 
  out.coda  <- coda.samples(model, variable.names=model_parameters,n.iter=2*nSamples) 
  save(out.coda,file = "out.coda_JAGS_VAR1.RData")
}else{
  load("out.coda_JAGS_VAR1.RData")
}
#plot(out.coda )

if(n.chains > 1)
{
  gelman.srf <-gelman.diag(out.coda)
  plot(gelman.srf$psrf,main = "Gelman Diagnostic")
}

chains.ess <- lapply(out.coda,effectiveSize)
first.chain.ess <- chains.ess[1]
plot(unlist(first.chain.ess), main="Effective Sample Size")

chain <- out.coda[[1]]
posterior.means <- list()
posterior.modes <- list()
for( i in 1:length(colnames(chain)) )
{  
  colname <- colnames(chain)[i]
  samples <- chain[,i]
  posterior.means[colname] <-mean(samples)
  posterior.modes[colname] <-mlv(samples)$M
}

# plot(posterior.means,posterior.modes) # Nice and unimodal

Sigma.map <- unlist(posterior.means)[1:36]
Sigma.model<- unlist(sig$mat)
plot(Sigma.map,Sigma.model)
rho.map <- unlist(posterior.means)[43]

nu.map <-  unlist(posterior.means)[37:42]
 
# y.map <-  matrix(unlist(posterior.means)[44:2233],ncol=6, byrow=FALSE)
# plot(y[,1]-y.map[,1]) 
```

# Missing data imputation in Openbugs - a simple generative model of $\theta$ with no serial correlation accounted for

Here we try to implement missing data imputation in OpenBugs. We use the data set y1 and fit at multivariate model to the data

$$\theta[t] \sim N(Y1[t], \Sigma)$$ we use BUGS and are imputing the missing data in Y1.

Notes :
- This does not take into account the temporal correlation 
- Openbugs was VERY HARD TO FIND on the internet - their website is down.
- We can not yet port the VAR(1) model toOpenBugs for further refinement.  

```{r, results='hide', message=FALSE, warning=FALSE}
rm(list = ls())
setwd("d:/brucebcampbell-git/bayesian-learning-with-R")
load("E2.RData")
library(R2OpenBUGS)
library(rjags)
library(coda)
library(modeest)
N <- nrow(Y1)
p = 6
x <- scale(Y1)

mlr_model2 <- function(){
  for(i in 1:N)
  {
    theta[i,1:p] ~ dmnorm(x[i,1:p] ,precision2[,])
  }
  
  # Prior for likelihood parameters: mu2, precision2, rho
  rho  ~  dunif(-1,1)
  
  for(j in 1:p)
  {
    mu2[j]  ~  dnorm(0,0.01)
  }
  
  precision2[1:p,1:p] ~ dwish(R[,],k)
  
  # Missing data model for x
  for(i in 1:N){
    x[i,1:p]~dmnorm(x_mn[],x_prec[,])
  }
  
  # Priors for missing-data model parameters
  for(j in 1:p){
    x_mn[j]~dnorm(0,0.01)
  }
  x_prec[1:p,1:p]~dwish(R[,],k)
  x_cov[1:p,1:p]<-inverse(x_prec[,])
  
  k <- p+0.1
  for(j1 in 1:p)
  {
    for(j2 in 1:p)
    {
      R[j1,j2] <- 0.1*equals(j1,j2)
    }
  }
}

n.chains = 2
nSamples=10000
stacks_dat <- list(x=x,p = 6,   N = 365)
mlr_inits <- function() {   list( rho = 0.00) }

if(FALSE){
  samps <- bugs(data = stacks_dat, 
              inits = mlr_inits, 
              parameters.to.save = c("theta"), 
              model.file = mlr_model2, 
              codaPkg = TRUE,
              n.chains = n.chains, n.burnin=2000, n.iter = nSamples, n.thin=10, DIC=F)

out.coda <- read.bugs(samps)
save(out.coda,file = "out.coda_BUGS_MVN.RData")
}else{
  load("out.coda_BUGS_MVN.RData")
}

if(n.chains > 1)
{
  gelman.srf <-gelman.diag(out.coda)
  count.coeff.gt <- sum(gelman.srf$psrf>1.1)
  count.coeff.gt
}

chains.ess <- lapply(out.coda,effectiveSize)

first.chain.ess <- chains.ess[1]
plot(unlist(first.chain.ess), main="Effective Sample Size")

chain <- out.coda[[1]]
posterior.means <- list()
posterior.modes <- list()

for( i in 1:(365*6) )
{  
  colname <- colnames(chain)[i]
  
  samples <- chain[,i]
  
  posterior.means[i] <-mean(samples)
  
  posterior.modes[i] <-mlv(samples)$M
}

plot(posterior.means, posterior.modes)

theta.map <-  matrix(unlist(posterior.means)[1:2190],ncol=6, byrow=FALSE)

unscaled.theta.map <- ( theta.map +colMeans(Y1,na.rm = TRUE)) *  apply(Y1, 2,sd,na.rm = TRUE)

write.csv(unscaled.theta.map,file = "unscaled-theta-map.csv")
hist(unscaled.theta.map[,1])

hist(Y1[,1])
```


# VAR(1) on Y1 with imputation in OpenBugs

Thank goodness for StackOverflow.  I was stuck for two days trying to port the VAR model from JAGS to OpenBugs.
This SO posting helped, but it took far too long to get there. 
https://stackoverflow.com/questions/23968710/openbugs-error-expected-multivariate-node


```{r, results='hide', message=FALSE, warning=FALSE}
rm(list = ls())
setwd("d:/brucebcampbell-git/bayesian-learning-with-R")
load("E2.RData")
library(R2OpenBUGS)
library(rjags)
library(coda)
library(modeest)
N <- nrow(Y1)
p = 6

mlr_model2 <- function(){

      for(i in 2:N) {
    Y1pred[i,1:p] ~ dmnorm( theta[i,1:p] ,precision[,])
       for(j in 1:p){
         theta[i,j]<-  mu2[j] + rho * Y1[i-1,j]
       }
  }
  
  # Priors
  rho  ~  dunif(-1,1)
  
  for(j in 1:p)
  {
    mu2[j]  ~  dnorm(0,0.01)
  }
  
  precision[1:p,1:p] ~ dwish(R[,],k)
  
  # Missing data model for Y1
  for(i in 1:N){
    Y1[i,1:p]~dmnorm(Y1_mn[],Y1_prec[,])
  }
  
  # Priors for missing-data model parameters
  for(j in 1:p){
    Y1_mn[j]~dnorm(0,0.01)
  }
  Y1_prec[1:p,1:p]~dwish(R[,],k)
  Y1_cov[1:p,1:p]<-inverse(Y1_prec[,])
  
  k <- p+0.1
  for(j1 in 1:p)
  {
    for(j2 in 1:p)
    {
      R[j1,j2] <- 0.1*equals(j1,j2)
    }
  }
}
n.chains = 1
nSamples=10000
Y1.scaled <- scale(Y1)
stacks_dat <- list(Y1=Y1.scaled,p = 6,   N = 365)
mlr_inits <- function() {   list( rho = 0.00) }
parameters.to.save = c("Y1pred","precision")
if(TRUE)
{
  samps <- bugs(data = stacks_dat,inits = mlr_inits, parameters.to.save =parameters.to.save, 
              model.file = mlr_model2, 
              codaPkg = TRUE,
              n.chains = n.chains, n.burnin=1000, n.iter = nSamples, n.thin=10, DIC=F)

out.coda <- read.bugs(samps)
save(out.coda,file = "out.coda_BUGS_VAR.RData")

}else{
  load("out.coda_BUGS_MVN.RData")
}

if(n.chains > 1)
{
  gelman.srf <-gelman.diag(out.coda)
  count.coeff.gt <- sum(gelman.srf$psrf>1.1)
  count.coeff.gt
}

chains.ess <- lapply(out.coda,effectiveSize)

first.chain.ess <- chains.ess[1]
plot(unlist(first.chain.ess), main="Effective Sample Size")

chain <- out.coda[[1]]
posterior.means <- list()
posterior.modes <- list()

for( i in 1:(365*6) )
{  
  colname <- colnames(chain)[i]
  
  samples <- chain[,i]
  
  posterior.means[i] <-mean(samples)
  
  posterior.modes[i] <-mlv(samples)$M
}

plot(posterior.means, posterior.modes)

theta.map <-  matrix(unlist(posterior.means)[1:2190],ncol=6, byrow=FALSE)

unscaled.theta.map <- ( theta.map +colMeans(Y1,na.rm = TRUE)) *  apply(Y1, 2,sd,na.rm = TRUE)

write.csv(unscaled.theta.map,file = "unscaled-theta-map.csv")
hist(unscaled.theta.map[,1])

hist(Y1[,1])
```




# VAR(1) on Y1 with starting values, Y2 and Y3 incorporated imputation in OpenBugs

```{r, results='hide', message=FALSE, warning=FALSE}
rm(list = ls())
setwd("c:/e/brucebcampbell-git/bayesian-learning-with-R")
load("E2.RData")
library(R2OpenBUGS)
library(rjags)
library(coda)
library(modeest)
N <- nrow(Y1)
p = 6

mlr_model2 <- function(){

      for(i in 2:N) {
    Y1pred[i,1:p] ~ dmnorm( theta[i,1:p] ,precision[,])
        for(j in 1:p){
         theta[i,j]<-  mu2[j] + rho * Y1[i-1,j]
        }
        
        
        Y2[i] ~ dnorm( rtheta[i,1],sigmaY)
        
        rm[i] <-theta[i,1]
        #Y3[i] ~ dnorm( thetaMean[i],sigmaY3 )
        thetaMean[i] <- theta[i,1]/6 +theta[i,2]/6 +theta[i,3]/6 +theta[i,4]/6 +theta[i,5]/6 +theta[i,6]/6
       
        
  }
  
  # Priors
  rho  ~  dunif(-1,1)
  sigmaY  ~  dnorm(0,0.01)
    
  for(j in 1:p)
  {
    mu2[j]  ~  dnorm(0,0.01)
  }
  
  precision[1:p,1:p] ~ dwish(R[,],k)
  
  # Missing data model for Y1
  for(i in 1:N){
    Y1[i,1:p]~dmnorm(Y1_mn[],Y1_prec[,])
  }
  
  # Priors for missing-data model parameters
  for(j in 1:p){
    Y1_mn[j]~dnorm(0,0.01)
  }
  Y1_prec[1:p,1:p]~dwish(R[,],k)
  Y1_cov[1:p,1:p]<-inverse(Y1_prec[,])
  
  k <- p+0.1
  for(j1 in 1:p)
  {
    for(j2 in 1:p)
    {
      R[j1,j2] <- 0.1*equals(j1,j2)
    }
  }
}

Y1.scaled <- scale(Y1)
Y2.scaled <- scale(Y2)
Y3.scaled <- scale(Y3)
stacks_dat <- list(Y1=Y1.scaled,Y2=Y2.scaled, p = 6,   N = 365)
mlr_inits <- function() {   list( rho = 0.00) }
parameters.to.save = c("Y1pred","precision")
#if(TRUE)
#{
  samps <- bugs(data = stacks_dat,inits = mlr_inits, parameters.to.save =parameters.to.save, 
              model.file = mlr_model2, 
              codaPkg = TRUE,
              n.chains = 1, n.burnin=50, n.iter = 200, DIC=F)#, n.thin=10

```




```{r}
  
  
out.coda <- read.bugs(samps)
save(out.coda,file = "out.coda_BUGS_VAR.RData")
# 
# }else{
#   load("out.coda_BUGS_MVN.RData")
# }

if(n.chains > 1)
{
  gelman.srf <-gelman.diag(out.coda)
  count.coeff.gt <- sum(gelman.srf$psrf>1.1)
  count.coeff.gt
}

chains.ess <- lapply(out.coda,effectiveSize)

first.chain.ess <- chains.ess[1]
plot(unlist(first.chain.ess), main="Effective Sample Size")

chain <- out.coda[[1]]
posterior.means <- list()
posterior.modes <- list()

for( i in 1:(365*6) )
{  
  colname <- colnames(chain)[i]
  
  samples <- chain[,i]
  
  posterior.means[i] <-mean(samples)
  
  posterior.modes[i] <-mlv(samples)$M
}

plot(posterior.means, posterior.modes)

theta.map <-  matrix(unlist(posterior.means)[1:2190],ncol=6, byrow=FALSE)

unscaled.theta.map <- ( theta.map +colMeans(Y1,na.rm = TRUE)) *  apply(Y1, 2,sd,na.rm = TRUE)

write.csv(unscaled.theta.map,file = "unscaled-theta-map.csv")
hist(unscaled.theta.map[,1])

hist(Y1[,1])
```
